---
title: "FiveThirtyEight Data"
author: "Adeel Abbasi"
date: "September 17, 2018"
output: html_document
---


## Github Cloning Step Question # 1

### Step 1 - Create a directory
```{r}
# mkdir Caesar
```

### Step 2 - Navigate
```{r}
# cd Caesar
```

### Step 3 - Clone a directory 
```{r}
# git clone https://github.com/caesar0301/awesome-public-datasets
```

## Data  - Question # 2

### Step A

In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.

#### Read the file
a.	In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df.  This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status

```{r}

url <- "awesome-public-datasets/Datasets/titanic.csv"
df = read.csv(file = url)


```

#### Count the male and feamle in the dataframe

```{r}

summary(df$Sex)

```

#### Generating a frequency of male/female
b. Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males.  Be sure to give an accurate title and label the axes.

```{r}

plot(df$Sex, ylab = 'Frequency', main = 'Titanic Data' )

```

### Find mean for Age, Fare and Survived
c.	Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival.  Make sure the output is a real number for all three means.

```{r}
sapply(df[, c('Age', 'Fare', 'Survived')], function(x) round(mean(x, na.rm = TRUE), 2))
# sapply(df[, c('Age', 'Fare', 'Survived')], function(x) mean(x, na.rm = TRUE))
# sapply(df[, c('Age', 'Fare', 'Survived')], mean, na.rm = TRUE)


```

## Function Building - Question # 3

```{r}

helper.loaddata <- function(filepath){
  sleepdata = read.csv(filepath)
    
  Age.Median <- median(sleepdata$Age, na.rm = TRUE)
  Sleep.Minimum <- min(sleepdata$Duration, na.rm = TRUE)
  Sleep.Maximum <- max(sleepdata$Duration, na.rm = TRUE)
  RSES.Mean <- mean(sleepdata$RSES, na.rm = TRUE)
  RSES.SD <- sd(sleepdata$RSES, na.rm = TRUE)
  
  report <- data.frame(Age.Median / 5, RSES.Mean/5, RSES.SD/5,
              Sleep.Maximum - Sleep.Minimum)
  
  colnames(report) <- c('MedianAge', 'SelfEsteem', 'SE_SD', 'DurationRange')
  report <- round(report, 2)
  
  report
}

url <- 'sleep_data_01.csv'
sleepdata = helper.loaddata(url)
sleepdata


```


### FiveThirtyEight Data - Question # 4

```{r}
# load the fivethirdyeight dataset collection
library("fivethirtyeight")

# load the package into a variable to find a dataset name
pkg <- data(package = "fivethirtyeight")

# identify the 22nd dataset name
dataset.name <- pkg$results[22, "Item"]

# print the datasetname 
paste("22nd Dataset name is", dataset.name)

#get the URL of the article 
datasets <- read.csv("https://goo.gl/OT8iHa")

url = subset(datasets, DATAFRAME_NAME==dataset.name)
paste('News Story URL is', url$ARTICLE_URL)

```

#### Dimensions of the data frame
```{r}
# hold the dataset into a df variable
df <- get(dataset.name)

dim(df)

```

#### columns name of the data frame
```{r}
names(df)

```

#### Get Columns names count of the cols - Question # 5
```{r}

# Column names 
colnames(df)

# Number of columns
dim(df)[2]


```

#### Generate a count of each unique major_category 
```{r}

library("sqldf")
major_count <- sqldf("SELECT major_category,COUNT(*) as num_major_category FROM df GROUP BY major_category")

names(major_count) <- c('Major Category', 'Count')
major_count

```

#### barplot of major_count
```{r}
par(las=2)

major_count <- table(df[,'major_category'])

barplot(major_count, main = 'Major Categories', xlab = 'Frequency', col = heat.colors(12), density = 99, border = "dark blue", horiz = TRUE, space = 0.2)

```


#### Write data to CSV without header
```{r}

write.csv(df, file = 'fivethirtyeight.csv', row.names = FALSE)


```

#### Question # 6

https://github.com/AdeelAbbasi/SMUMSDShomeworkAssignment3

